---
title: "Assignment_6"
format: pdf
---

# Assignment 6 - Sobol and Dynamic Model

1.  Implement this model in R (as a differential equation).

```{r}
#| message: false
#| warning: false
# Load necessary packages
library(tidyverse)
library(deSolve)
library(here)
library(sensitivity)
```

```{r}
source("R/dforestgrowth.R")
```

2.  Run the model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C, and using the following parameters:
    2.  canopy closure threshold of 50 kgC
    3.  K = 250 kg C (carrying capacity)
    4.  r = 0.01 (exponential growth rate before before canopy closure)
    5.  g = 2 kg/year (linear growth rate after canopy closure)

```{r}
# Initial forest size
C_initial <- 10

# 300 year time span
simtimes <- seq(from = 1, to = 300, by = 1)

# Base params 
parms <- list(
  r  = 0.01,
  g = 2,
  canopy_thresh = 50,
  K = 250
)

# Ode solver
result_forest <- ode(y = C_initial, times = simtimes, func = dforestgrowth, parms = parms)

# Create data frame 
forest_df <- data.frame(time = result_forest[,1], C = result_forest[,2])
```

3.  Graph the results. Here you are graphing the trajectory with the parameters as given (e.g no uncertainty)

```{r, fig.width = 10, fig.height=5}
forest_df %>%
  # Plot x as year and y as forest size (C)
ggplot(aes(x = time, y = C)) +
  
  # show trajectory of growth
  geom_line(color = "forestgreen", linewidth = 1) +
  
  # we can highlight the canopy threshold at 50 kgc 
  geom_hline(aes(yintercept = 50, 
                 linetype = "Canopy Closure Threshold (50 kgC)"), 
             color = "orange") +
  
  # we can also highlight where the carrying capacity is (250 kgC)
  geom_hline(aes(yintercept = 250, 
                 linetype = "Carrying Capacity (250 kgC)"), 
             color = "brown") +

  labs(title = "Forest Growth Over 300 Years",
       x = "Year",
       y = "Forest Size (kg Carbon)") +
  theme_minimal()
```

4.  Run a sobol global (vary all parameters at the same time) sensitivity analysis that explores how the estimated maximum forest size (e.g maximum of 300 years, varies with these parameters:
    4.  pre canopy closure growth rate
    5.  post-canopy closure growth rate
    6.  canopy closure threshold and carrying capacity

Assume that parameters are all normally distributed with means as given above and standard deviation of 10% of mean value.

```{r}
# Set seed for reproducibility
set.seed(123)  

# Set num of samples 
np <- 200     

# Define sample set 1 (X1)
# Each parameter is norm dis (mean = base value and sd = 10% of mean)
X1 <- cbind.data.frame(
  r = rnorm(n = np, mean = 0.01, sd = 0.001),  
  g = rnorm(n = np, mean = 2,    sd = 0.2),    
  canopy_thresh = rnorm(n = np, mean = 50, sd = 5), 
  K = rnorm(n = np, mean = 250,  sd = 25) 
)

# Define sample set 2 (X2)
X2 <- cbind.data.frame(
  r  = rnorm(n = np, mean = 0.01, sd = 0.001),
  g = rnorm(n = np, mean = 2,    sd = 0.2),
  canopy_thresh = rnorm(n = np, mean = 50,   sd = 5),
  K= rnorm(n = np, mean = 250,  sd = 25)
)


# Create global sobol with parameters and samples
sens_forest <- sobolSalt(model = NULL, X1, X2, nboot = 300)

# Label columns
colnames(sens_forest$X) <- c("r", "g", "canopy_thresh", "K")

head(sens_forest$X)  
```

5\. Graph the results of the sensitivity analysis as a box plot of maximum forest size and record the two Sobol indices (S and T).

```{r}
# We have to build a wrapper function such that it runs ode 
# AND returns max forest size for just one parameter set
forest_wrapper <- function(r, g, canopy_thresh, K, C_initial, simtimes, odefunc) {
  # Group parameters into a list to pass into ode()
  parms <- list(r = r, g = g, canopy_thresh = canopy_thresh, K = K)
  
  # Run ode with defined params
  result <- ode(y = C_initial, times = simtimes, func = odefunc, parms = parms)
  
  # Create data frame by extracting first and second columns
  result_df <- data.frame(time = result[,1], C = result[,2])
  
  # Return max forest size (our metric)
  return(list(maxC = max(result_df$C)))
}

# Test on one row first
forest_wrapper(r = 0.01, g = 2, canopy_thresh = 50, K = 250,
               C_initial = 10, simtimes = simtimes, odefunc = dforestgrowth)
```

```{r}
# Run for all parameter sets
allforests <- as.data.frame(sens_forest$X) %>%
  pmap(forest_wrapper,
       C_initial = 10,
       simtimes  = simtimes,
       odefunc   = dforestgrowth)

# Extract max forest size into data frame
allres_max <- allforests %>% map_dfr(`[`, c("maxC"))
```

```{r}
# Graph metric (max forest size)
allres_max %>% 
  ggplot(aes(x = "", y = maxC)) +
  geom_boxplot(fill = "forestgreen", alpha = 0.5) +
  labs(title = "Sobol Sensitivity of Maximum Forest Size over 300 Years",
       x = "",
       y = "Maximum Forest Size (kg Carbon)") +
  theme_minimal()
```

```{r}
# Inform sobol of maximum forest sizes
sens_forest_maxC <- sensitivity::tell(sens_forest, allres$maxC)

# S indices (first order)
rownames(sens_forest_maxC$S) <- c("r", "g", "canopy_thresh", "K")
sens_forest_maxC$S

# T (total effect)
rownames(sens_forest_maxC$T) <- c("r", "g", "canopy_thresh", "K")
sens_forest_maxC$T
```

6\. In 2-3 sentences, discuss what the results of your simulation might mean. (For example think about how what parameters climate change might influence).

The results indicate that carrying capacity (K) and pre-canopy growth rate (*r*) are the most important drivers of maximum forest size, with K having the strongest influence. Post-canopy growth rate (*g*) has a moderate effect while the canopy cover while the canopy closure threshold contributes very little to the maximum forest size. Under climate change, early growth may increase due to warming and carbon dioxide fertilization; however, carrying capacity could decline from drought and disturbance, ultimately limiting longer term carbon storage.
