---
title: "Sobol Sensitivity Analysis Assignment 5"
format: pdf
author: "Vedika Shirtekar"
date: "2/10/2026"
---

# Sobol Sensitivity Analysis Assignment 5

## Part 1: Paper Discussion

**Selected paper: "**Sensitivity analysis of cooling demand applied to a large office building" by Zeferina et al. (2021)

In the study conducted by Zeferina et al. (2021), a Sobol sensitivity analysis revealed that ventilation rate is the most influential parameter for determining both annual and peak cooling demand in large office buildings, contributing to about 50-70% of the uncertainty in HVAC electricity use. This finding is critical for environmental management because it identifies a specific area where building designers and engineers can reduce the problem of oversizing cooling systems; this "system oversizing" problem could lead to inefficient energy use and unnecessary equipment costs. By focusing on more precise control and measurement of ventilation rates, building managers can better predict peak electricity loads, improving the resilience of power networks and reducing the overall carbon footprint of cooling-dominated office buildings.

## Part 2: Atmospheric Conductance and Sobol

#### Import the necessary libraries

```{r}
#| warning: false
#| message: false
library(sensitivity)
library(tidyverse)
library(purrr)
library(here)
```

#### Define parameter distributions

```{r}
np <- 1000
set.seed(123) # For reproducibility

## Create randomized distributions for constructing X1
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.1, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.1, n = np)
v <- rnorm(mean = 3, sd = .5, n = np)
height <- runif(min = 9.5, max = 10.5, n = np)

X1 <- cbind.data.frame(k_o, k_d, v, height = height)

## Repeat for X2
k_o <- rnorm(mean = 0.1, sd = 0.1 * 0.1, n = np)
k_d <- rnorm(mean = 0.7, sd = 0.7 * 0.1, n = np)
v <- rnorm(mean = 3, sd = .5, n = np)
height <- runif(min = 9.5, max = 10.5, n = np)
X2 <- cbind.data.frame(k_o, k_d, v, height = height)
```

#### Generate Sobol parameter matrix

```{r}
# Use sobolSalt for generating matrix of parameter combos 
# to test (recall that we want to see how total variation occurs 
# in the output of the model when one parameter is experimented with)
source(here("R/Catm.R"))
sens_Catm_Sobol <- sobolSalt(model = NULL, X1, X2, nboot = 100)
summary(sens_Catm_Sobol) # Should be empty
```

#### Run model on parameter sets

```{r}
parms <- as.data.frame(sens_Catm_Sobol$X)
colnames(parms) <- colnames(X1)
#colnames(parms) <- c("v", "height", "k_d", "k_o")
res <- pmap_dbl(parms, Catm) # Run Catm function for every row
```

#### Calculate Sobol indices

Here, we feed the results from each row based on the `Catm` function into the sensitivity object.

```{r}
sens_Catm_Sobol <- sensitivity::tell(sens_Catm_Sobol, res, res.names = "ga")

# Add column names
row.names(sens_Catm_Sobol$S) <- colnames(parms)
row.names(sens_Catm_Sobol$T) <- colnames(parms)
```

```{r}
## Take a look at the results
print("First order sensitivity:")
sens_Catm_Sobol$S
```

```{r}
## Take a look at the results
print(paste("Total effects:"))
sens_Catm_Sobol$T
```

#### Visualization and Analysis

Plot the following:

-   uncertainty in the output

-   parameter-output variation for the most sensitive parameters

```{r}
## Combine the parameters with their associated results
param_results <- cbind.data.frame(parms, gs = sens_Catm_Sobol$y)
```

```{r}
## Plot uncertainty 
param_results %>% 
  ggplot(aes(x = gs)) + 
  geom_histogram(fill = "skyblue", color = "white", bins = 30) +
  labs(title = "Uncertainty in Atmospheric Conductance",
       x = "Conductance (mm/s)", 
       y = "Frequency") + 
  geom_vline(xintercept = mean(param_results$gs), col = "red") + 
  theme_light()
```

```{r}
## Plot parameter-output variation for the most sensitive parameters

# First sensitive parameter is k_d
param_results %>% 
ggplot(aes(x = v, y = gs, color = k_d)) + 
  geom_point(alpha = 0.5) + 
  scale_color_viridis_c() + 
  labs(title = "Sensitivity of Conductance to Windspeed",
       subtitle = "Colored by k_d (most important parameter)",
       x = "Windspeed (m/s)",
       y = "Conductance (mm/s)",
       color = "k_d") +
  theme_light()

# Second sensitive parameter besides windspeed
param_results %>% 
ggplot(aes(x = v, y = gs, color = k_o)) + 
  geom_point(alpha = 0.5) + 
  scale_color_viridis_c() + 
  labs(title = "Sensitivity of Conductance to Windspeed",
       subtitle = "Colored by k_o (second most important parameter excluding windspeed)",
       x = "Windspeed (m/s)",
       y = "Conductance (mm/s)",
       color = "k_o") +
  theme_light()
```

#### Discussion 

The Sobol sensitivity analysis total effect indices in this assignment reveal that the model's sensitivity to environmental forcing has increased significantly in this new environmental setting characterized by higher (and more variable) windspeeds and shorter vegetation. In our class example, windspeed is a relatively minor contributor to the atmospheric conductance variance (T = 0.19) with the constant `k_d` (T = 0.6) dominating the model. In this environmental setting defined for this assignment (higher, more variable windspeeds and shorter vegetation), the influence of windspeed on the atmospheric conductance is much larger (T = 0.35). The `k_d` parameter is still the most influential factor (T = 0.44) but its dominance is decreased under the new environmental settings. This difference indicates that in windier environments with shorter canopies/vegetation heights, meteorological conditions like windspeed may play a larger role in determining the uncertainty in atmospheric conductance than these conditions would do in a more sheltered environment dominated by tall vegetation, as seen in class.
