---
title: "Assignment_6"
format: html
---

1.  Implement this model in R (as a differential equation)

```{r}
# Load necessary packages
library(tidyverse)
library(deSolve)
library(here)
library(sensitivity)
```

```{r}
# Create deforest function to include differential equations in different stages of life cycle 
# T = time 
# C = forest size
# params are established (r, g, canopy_thresh, carrying capacity K)
dforestgrowth <- function(Time, C, parms) {
  # if forest still below threshold (50 kgC)
  if (C < parms$canopy_thresh) {
  # grow exponentially (r* current forest size)
    dC <- parms$r * C
   
  } else {
    # if forest is >= threshold 
    dC <- parms$g * (1 - C / parms$K)
    # linear growth towards K 
  }
  
  return(list(dC)) # output is a list according to deSolve
}
```

2.  Run the model for 300 years (using the ODE solver) starting with an initial forest size of 10 kg/C, and using the following parameters: canopy closure threshold of 50 kgC = 250 kg C (carrying capacity) = 0.01 (exponential growth rate before before canopy closure) = 2 kg/year (linear growth rate after canopy closure)

```{r}
# initial forest size
C_initial <- 10

# 300 year time span
simtimes <- seq(from = 1, to = 300, by = 1)

# base params 
parms <- list(
  r  = 0.01,
  g = 2,
  canopy_thresh = 50,
  K = 250
)

# ode solver
result <- ode(y = C_initial, times = simtimes, func = dforestgrowth, parms = parms)

# data frame 
result_df <- data.frame(time = result[,1], C = result[,2])
```

2.  Graph the results. Here you are graphing the trajectory with the parameters as given (e.g no uncertainty)

    ```{r, fig.width = 10, fig.height=5}
    result_df %>%
      # plot x as year and y as forest size (C)
    ggplot(aes(x = time, y = C)) +
      
      # show trajectory of growth
      geom_line(color = "forestgreen", linewidth = 1) +
      
      # we can highlight the canopy threshold at 50 kgc 
      geom_hline(aes(yintercept = 50, 
                     linetype = "Canopy Closure Threshold (50 kgC)"), 
                 color = "orange") +
      
      # we can also highlight where the carrying capacity is (250 kgC)
      geom_hline(aes(yintercept = 250, 
                     linetype = "Carrying Capacity (250 kgC)"), 
                 color = "brown") +

      labs(title = "Forest Growth Over 300 Years",
           x = "Year",
           y = "Forest Size (kg Carbon)") +
      theme_minimal()
    ```

3.  Run a sobol global (vary all parameters at the same time) sensitivity analysis that explores how the estimated maximum forest size (e.g maximum of 300 years, varies with these parameters:

    1.  pre canopy closure growth rate ( )
    2.  post-canopy closure growth rate ( )
    3.  canopy closure threshold and carrying capacity( )

```{r}
# set seed for reproducibility
set.seed(123)  
# set num of samples 
np <- 200     

# define sample set 1 (X1)
# each parameter is norm dis (mean = base value and sd = 10% of mean)
X1 <- cbind.data.frame(
  r = rnorm(n = np, mean = 0.01, sd = 0.001),  
  g = rnorm(n = np, mean = 2,    sd = 0.2),    
  canopy_thresh = rnorm(n = np, mean = 50, sd = 5), 
  K = rnorm(n = np, mean = 250,  sd = 25) 
)

# define sample set 2 (X2)
X2 <- cbind.data.frame(
  r  = rnorm(n = np, mean = 0.01, sd = 0.001),
  g = rnorm(n = np, mean = 2,    sd = 0.2),
  canopy_thresh = rnorm(n = np, mean = 50,   sd = 5),
  K= rnorm(n = np, mean = 250,  sd = 25)
)


# create global sobol with parameters and samples
sens_forest <- sobolSalt(model = NULL, X1, X2, nboot = 300)

# label columns
colnames(sens_forest$X) <- c("r", "g", "canopy_thresh", "K")

head(sens_forest$X)  
```

Assume that parameters are all normally distributed with means as given above and standard deviation of 10% of mean value

5\. Graph the results of the sensitivity analysis as a box plot of maximum forest size and record the two Sobol indices (S and T).

```{r}
# we have to build a wrapper function such that it runs ode AND returns max forest size for just one parameter set
forest_wrapper <- function(r, g, canopy_thresh, K, C_initial, simtimes, odefunc) {
  # group parameters into a list to pass into ode()
  parms <- list(r = r, g = g, canopy_thresh = canopy_thresh, K = K)
  
  # run ode with defined params
  result <- ode(y = C_initial, times = simtimes, func = odefunc, parms = parms)
  
  # create data frame by extracting first and second columns
  result_df <- data.frame(time = result[,1], C = result[,2])
  # convert to data frame
    
  # return max forest size (our metric)
  return(list(maxC = max(result_df$C)))
}

# test on one row first
forest_wrapper(r = 0.01, g = 2, canopy_thresh = 50, K = 250,
               C_initial = 10, simtimes = simtimes, odefunc = dforestgrowth)
```

```{r}
# run for all parameter sets
allresults <- as.data.frame(sens_forest$X) %>%
  pmap(forest_wrapper,
       C_initial = 10,
       simtimes  = simtimes,
       odefunc   = dforestgrowth)

# extract max forest size into data frame
allres <- allresults %>% map_dfr(`[`, c("maxC"))
```

```{r}
# graph metric (max forest size)
# boxplot
allres %>% 
  ggplot(aes(x = "", y = maxC)) +
  geom_boxplot(fill = "forestgreen", alpha = 0.5) +
  labs(title = "Sobol Sensitivity of Maximum Forest Size over 300 Years",
       x = "",
       y = "Maximum Forest Size (kg Carbon)") +
  theme_minimal()
```

```{r}
# inform sobol of maximum forest sizes
sens_forest_maxC <- sensitivity::tell(sens_forest, allres$maxC)

# S indices (first order)
rownames(sens_forest_maxC$S) <- c("r", "g", "canopy_thresh", "K")
sens_forest_maxC$S

# T (total effect)
rownames(sens_forest_maxC$T) <- c("r", "g", "canopy_thresh", "K")
sens_forest_maxC$T
```

6\. In 2-3 sentences, discuss what the results of your simulation might mean. (For example think about how what parameters climate change might influence).
